---
title: "pcirm-examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pcirm-examples}
  %\VignetteEncoding{UTF-8}
  <!-- \usepackage[utf8]{inputenc} -->
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "R>"
)
```

**Note: the estimation process can be time consuming depending on the computing power. You can same some time by reducing the length of the chains.**

## Dichotomous Data with Local Dependence and Minor Trait:

1) Load the package, obtain the data, loading pattern (*qlam*), and no. of factors/traits.

```{r}
library(LAWBL)
dat <- sim24ccfa21$dat
head(dat)
J <- ncol(dat) # no. of items
qlam <- sim24ccfa21$qlam
qlam
K <- ncol(qlam) # no. of factors

```

2) PCIRM with primary loadings and three traits specified. Longer chain is suggested for stabler performance (*burn=iter=5,000* by default).

```{r}
ipf <- 8
Q<-matrix(-1,J,K-1); # -1 for unspecified items
for (k in 1:(K-1)){
  tmp<-(k-1)*ipf
  Q[(tmp+1):(tmp+ipf),k]<-1
}
Q
m0<-pcirm(dat = dat,Q=  Q,LD = TRUE,cati = -1,burn = 2000,iter = 2000,verbose = TRUE)
summary(m0)
summary(m0, what = 'qlambda')
summary(m0, what = 'offpsx') #summarize significant LD terms
summary(m0,what='int')
summary(m0,what='eigen')

#plotting factorial eigenvalue
plot_eigen(m0) # trace
plot_eigen(m0, what='density') #density
plot_eigen(m0, what='APSR') #adj, PSRF
```

3) PCIRM-LI with four traits and two cross-loadings specified based on results from previous step. Longer chain is suggested for stabler performance.

```{r eval=FALSE}
Q<-cbind(Q,-1);
Q[c(15:16),K] <- 1

m1<-pcirm(dat = dat,Q=  Q,LD = FALSE, cati = -1,burn = 4000,iter = 4000,verbose = TRUE)
summary(m1)
summary(m1, what = 'qlambda') #close to qlam
```

4) PCIRM with four traits and four cross-loadings specified based on results from previous step. Longer chain is suggested for stabler performance.

```{r eval=FALSE}
tmp<-summary(m1, what="qlambda")
Q<-matrix(-1,J,K)
Q[tmp!=0]<-1
Q

m2<-pcirm(dat = dat,Q=  Q,LD = TRUE,cati = -1,burn = 4000,iter = 4000,verbose = TRUE)
summary(m2)
summary(m2, what = 'qlambda')
summary(m2, what = 'offpsx') #summarize significant LD terms
summary(m2,what='int')
summary(m2,what='eigen')

#plotting factorial eigenvalue
plot_eigen(m2) # trace
plot_eigen(m2, what='density') #density
plot_eigen(m2, what='APSR') #adj, PSRF
```

